Supporting information: Copulas and their potential for ecology

### 2019-09-29 Paper.Rmd
table fixed as suggested by DCR

dt <- tibble(Description = c("1. p-value, independence test",
           "2. Minimum-AIC copula",
           "3. AIC weight of minimum-AIC copula",
           "4. AIC weight of normal copula",
           "5. p-value, Cramer-von Mises goodness of fit test",
           "6. p-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates", # this row will be commented out, that's why the numbering starts from 10 in the next line
           "10. Placement of cor$_{FS}$-cor$_{LS}$ in surrogate distribution (see caption)")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
tab_bivar_summary<-tab_bivar_summary[c(1:9,11),]

knitr::kable(tab_bivar_summary, 
             format="latex", escape=F,row.names=F,
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values (rows 5-6) are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. Row 10 shows values as in Table \\ref{tab_soilCN_asym}, upper right table entry. Although the result shown in row 10 for the soil C and N data is non-significant, see Table \\ref{tab_soilCN_asym} for significant results using the $P$ and $D^2$ statistics. The first 3 datasets use $(l_b,u_b)=(0,0.1)$ for the first slice (FS) and $(0.9,1)$ for last slice (LS), whereas $(0, 0.25)$ and $(0.75,1)$ were used for Cedar Creek due to fewer data for that system. Results for Cedar Creek for the year 2000 are shown. See Table \\ref{tab_soilCNfit} for copula family abbreviations.\\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
             )%>%column_spec(2:5,width="1.4cm")%>%
             column_spec(1, width = "8.5cm")%>%kable_styling(latex_options = "basic",position = "center")
```

@@

tab_multivar_summary<-rbind(tab_multivar_summary[c(1:6),],
                            two_tab_in1[1:2,], #CI of LT, CI of UT
                            tab_multivar_summary[11,],
                            two_tab_in1[5,] #CI of Pl-Pu
                           )

row.names(tab_multivar_summary)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(
            Description = c("1. Location pairs (excluding self comparisons)",
           "2. Number of non-independent pairs",
           "3. Number of pairs with non-normal copula as best fit",
           "4. Average AIC-weights of best-fitting copulas across location pairs",
           "5. Average AIC-weights of normal copula across pairs",
           "6. Number of well-fitting location pairs",
           "7. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. lower-tail dependence",
           "8. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. upper-tail dependence",
           "9. Percent pairs showing stronger lower tail dependence",
           "10. 2.5$^{th}$ and 97.5$^{th}$ quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
           )

@@

knitr::kable(tab_multivar_summary[1:9,], 
             format="latex", escape = F,
             caption.short = "Summary of results for multivariate datasets",
             caption = "Summary of Q1 results for multivariate datasets. Rows 3-9 of the table were computed for the non-independent pairs (row 2) only. A well-fitting location pair (row 6) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests. \\label{tab_multivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ","Green spruce aphid abundance",
                        "Leaf-curling plum aphid first flight",
                        "\\textit{Ceratium furca} abundance","Methane-flux")
             )%>%column_spec(2:5,width="2.5cm")%>%
             column_spec(1, width = "6cm")%>%kable_styling(latex_options = "basic",position = "center")
```

>>>>

dt <- tibble(Description = c("1. p-value, independence test",
           "2. Minimum-AIC bestfit copula",
           "3. AIC for bestfit copula",
           "4. AIC for normal copula",
           "5. p-value, Cramer-von Mises goodness of fit test",
           "6. p-value, Kolmogorov-Smirnov goodness of fit test",
           "7. Model averaged lower-tail dependence",
           "8. Model averaged upper-tail dependence",
           "9. Model averaged lower- minus upper-tail dependence",
           "10. Asymmetry rank in surrogates", # this row will be commented out, that's why the numbering starts from 10 in the next line
           "10. Placement of cor$_{FS}$-cor$_{LS}$ in surrogate distribution (see caption)")
)
tab_bivar_summary<-cbind(dt,tab_bivar_summary)
tab_bivar_summary<-tab_bivar_summary[c(1:9,11),]

knitr::kable(tab_bivar_summary, 
             format="latex", escape=F,row.names=F,
             caption = "Summary of Q1 results for bivariate datasets. The $p$-values (rows 5-6) are for the minimum-AIC copula. Model averaging used for rows 7-9 was based on AIC weights. Row 10 shows values as in Table \\ref{tab_soilCN_asym}, upper right table entry. Although the result shown in row 10 for the soil C and N data is non-significant, see Table \\ref{tab_soilCN_asym} for significant results using the $P$ and $D^2$ statistics. The first 3 datasets use $(l_b,u_b)=(0,0.1)$ for the first slice (FS) and $(0.9,1)$ for last slice (LS), whereas $(0, 0.25)$ and $(0.75,1)$ were used for Cedar Creek due to fewer data for that system. Results for Cedar Creek for the year 2000 are shown. See Table \\ref{tab_soilCNfit} for copula family abbreviations.\\label{tab_bivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ", "Soil C and N", "Bird masses and BMR", "Mammal masses and BMR", "Cedar Creek data")
             )%>%column_spec(2:5,width="1.5cm")%>%
             column_spec(1, width = "8.4cm")%>%kable_styling(latex_options = "basic",position = "center")
```

@@

tab_multivar_summary<-rbind(tab_multivar_summary[c(1:3,5,6),],
                            two_tab_in1[1:2,], #CI of LT, CI of UT
                            tab_multivar_summary[11,],
                            two_tab_in1[5,] #CI of Pl-Pu
                           )

row.names(tab_multivar_summary)<-NULL
library(kableExtra)
library(dplyr)
library(tibble)
dt <- tibble(
            Description = c("1. Location pairs (excluding self comparisons)",
           "2. Number of non-independent pairs",
           "3. Number of pairs with non-normal copula as best fit",
           #"4. Avg. of normal-AICw minus bestfit-copula AICw across location pairs",
           "4. Average of AIC differences (normal minus bestfit-copula) across location pairs",
           "5. Number of well-fitting location pairs",
           "6. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. lower-tail dependence",
           "7. 2.5$^{th}$ and 97.5$^{th}$ quantiles of model-avg. upper-tail dependence",
           "8. Percent pairs showing stronger lower tail dependence",
           "9. 2.5$^{th}$ and 97.5$^{th}$ quantiles of $<$P$_{l}$ - P${_u}$$>$ across all location pair")
           )

@@

knitr::kable(tab_multivar_summary[1:8,], 
             format="latex", escape = F,
             caption.short = "Summary of results for multivariate datasets",
             caption = "Summary of Q1 results for multivariate datasets. Rows 3-8 of the table were computed for the non-independent pairs (row 2) only. A well-fitting location pair (row 5) was one for which the best-fitting copula had $p$-values $>0.01$ for both the Cramer-von Mises and Kolmogorov-Smirnov goodness of fit tests. \\label{tab_multivar_summary}",
             booktabs=T, linesep = "\\addlinespace",
             col.names = c(" ","Green spruce aphid abundance",
                        "Leaf-curling plum aphid first flight",
                        "\\textit{Ceratium furca} abundance","Methane-flux")
             )%>%column_spec(2:5,width="2.4cm")%>%
             column_spec(1, width = "6cm")%>%kable_styling(latex_options = "basic",position = "center")
```

### 2019-09-29 SupportingInformation.Rmd
table fixed as suggested by DCR

  m<-matrix(NA,1,10)
  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AICw",
                 "Normal.copula.AICw","p.CvM","p.KS",
                 "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
                 "Rank.(Pl-Pu)")
  m<-as.data.frame(m)

  m$p.Independence.test<-round(r$IndepTestRes,3)
  m$minimumAIC.bestfit.copula<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
  m$Bestfit.copula.AICw<-round(max(r$InfCritRes$AICw),3)
  m$Normal.copula.AICw<-round(r$InfCritRes$AICw[1],3)
  m$p.CvM<-r$GofRes_CvM
  m$p.KS<-r$GofRes_KS
  m$LT.model.avg<-round(r$relLTdep_AICw,3)
  m$UT.model.avg<-round(r$relUTdep_AICw,3)
  m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
  m$`Rank.(Pl-Pu)`<-rs$Rank_PlmPu_K
  mt<-as.data.frame(t(m))
  return(mt)

@@

<!--table for AIC-weights of normal copula used for fitting, green spruce aphid abundance data-->
```{r tab_count_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_count),
             format='latex',linesep = "",digits=3,
             caption.short = "AIC weights, normal copulas, green spruce aphid abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}. \\label{tab_count_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--table for AIC-weights of normal copula used for fitting, leaf plum curling aphid first flight data-->
```{r tab_ff_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
knitr::kable(normal_cop_AICw_matrix(r=RES_aphid_ff),
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, normal copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--table for AIC-weights of normal copula used for fitting, Ceratium furca abundance data-->
```{r tab_pns_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_plankton_north_sea)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, normal copulas, \\textit{Ceratium furca} abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
             booktabs=TRUE)
```

@@

<!--table for AIC-weights of normal copula used for fitting, methane-flux data-->
```{r tab_methane_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_methane)
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, normal copulas, methane-flux data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
             booktabs=TRUE)
```

@@

  m<-matrix(NA,1,17)
  colnames(m)<-c("num.pairs","num.non-indep","num.non-N",
                 "Avg.best.AICw","Avg.N.AICw","num.good.fit",
                 "0.025-CI.LT","0.975-CI.LT","0.025-CI.UT","0.975-CI.UT",
                 "percent.LT-UT>0",
                 "0.025-CI.Pl","0.975-CI-Pl","0.025-CI.Pu","0.975-CI.Pu","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu")
  m<-as.data.frame(m)

  normal_AICw<-normal_cop_AICw_matrix(r=r)

  d<-dim(r$gfc_numBS)[1]
  m$num.pairs<-(d^2)-d 
  m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)
  bcop<-r$info_ord_copcode[,,1] # matrix with best fit copula name
  bcop[bcop==Inf]<-NaN # replacing Independent pairs with NaN
  m$`num.non-N`<-sum(bcop!=1,na.rm = T)
  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN
  m$Avg.best.AICw<-round(mean(r$info_ord_AICw[,,1],na.rm = T),3)
  normal_AICw[normal_AICw==Inf]<-NaN
  m$Avg.N.AICw<-round(mean(normal_AICw,na.rm=T),3)
  pcvm<-r$gfc_p_CvM
  pcvm[pcvm==Inf]<-NaN
  pks<-r$gfc_p_KS
  pks[pks==Inf]<-NaN
  temp<-which((pcvm>0.01) & (pks>0.01),arr.ind=T)
  m$num.good.fit<-dim(temp)[1]
  t<-r$LTdep_AICw[is.finite(r$LTdep_AICw)]
  CI_LT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$`0.025-CI.LT`<-round(CI_LT[1],3)
  m$`0.975-CI.LT`<-round(CI_LT[2],3)
  t<-r$UTdep_AICw[is.finite(r$UTdep_AICw)]
  CI_UT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$`0.025-CI.UT`<-round(CI_UT[1],3)
  m$`0.975-CI.UT`<-round(CI_UT[2],3)
  m$`percent.LT-UT>0`<-round((sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)/((d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)))*100,2)
  m$`0.025-CI.Pl`<-round(rs$numericdf[5,3],3)
  m$`0.975-CI-Pl`<-round(rs$numericdf[5,4],3)
  m$`0.025-CI.Pu`<-round(rs$numericdf[6,3],3)
  m$`0.975-CI.Pu`<-round(rs$numericdf[6,4],3)
  m$`0.025-CI.Pl-Pu`<-round(rs$numericdf[10,3],3)
  m$`0.975-CI.Pl-Pu`<-list(round(rs$numericdf[10,4],3))
  #m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
  #m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
  #m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))

>>>>

  m<-matrix(NA,1,10)
  colnames(m)<-c("p.Independence.test","minimumAIC.bestfit.copula","Bestfit.copula.AIC",
                 "Normal.copula.AIC","p.CvM","p.KS",
                 "LT.model.avg","UT.model.avg","(LT-UT).model.avg",
                 "Rank.(Pl-Pu)")
  m<-as.data.frame(m)

  m$p.Independence.test<-round(r$IndepTestRes,3)
  m$minimumAIC.bestfit.copula<-r$InfCritRes$copname[which.min(r$InfCritRes$AIC)]
  m$Bestfit.copula.AIC<-round(min(r$InfCritRes$AIC),2)
  m$Normal.copula.AIC<-round(r$InfCritRes$AIC[1],2)
  m$p.CvM<-r$GofRes_CvM
  m$p.KS<-r$GofRes_KS
  m$LT.model.avg<-round(r$relLTdep_AICw,4)
  m$UT.model.avg<-round(r$relUTdep_AICw,4)
  m$`(LT-UT).model.avg`<-m$LT.model.avg-m$UT.model.avg
  m$`Rank.(Pl-Pu)`<-rs$Rank_PlmPu_K
  mt<-as.data.frame(t(m))
  return(mt)

@@

<!--table for AIC-weights of normal copula used for fitting, green spruce aphid abundance data-->
```{r tab_count_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_aphid_count)
h<-h$normal_cop_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format='latex',linesep = "",digits=3,
             caption.short = "AIC weights, normal copulas, green spruce aphid abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for green spruce aphid abundance data, for comparison with Table \\ref{tab_count_AICw}. \\label{tab_count_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--table for AIC-weights of normal copula used for fitting, leaf plum curling aphid first flight data-->
```{r tab_ff_AICw_normal, echo=F, results='markup'}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_aphid_ff)
h<-h$normal_cop_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format='latex',linesep = "",digits=3,
             caption.short="AIC weights, normal copulas, leaf-curling plum aphid first flight data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for leaf-curlung plum aphid first flight data, for comparison with Table \\ref{tab_ff_AICw}.\\label{tab_ff_AICw_normal}",
             booktabs=TRUE
             )
```

@@

<!--table for AIC-weights of normal copula used for fitting, Ceratium furca abundance data-->
```{r tab_pns_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_plankton_north_sea)
h<-h$normal_cop_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, normal copulas, \\textit{Ceratium furca} abundance data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for \\textit{Ceratium furca} abundance data, for comparison with Table \\ref{tab_pns_AICw}.\\label{tab_pns_AICw_normal}",
             booktabs=TRUE)
```

@@

<!--table for AIC-weights of normal copula used for fitting, methane-flux data-->
```{r tab_methane_AICw_normal,echo=F,results="markup"}
source("normal_cop_AICw.R")
h<-normal_cop_AICw_matrix(r=RES_methane)
h<-h$normal_cop_AICw
h[is.infinite(h)]<-NA
knitr::kable(h,
             format="latex",linesep = "",
             digits=3,
             caption.short="AIC weights, normal copulas, methane-flux data",
             caption = "$\\text{AIC}_{\\text{w}}$ for the normal copula family for each location pair for methane-flux data, for comparison with Table \\ref{tab_methane_AICw}.\\label{tab_methane_AICw_normal}",
             booktabs=TRUE)
```

@@

  m<-matrix(NA,1,17)
  colnames(m)<-c("num.pairs","num.non-indep","num.non-N",
                 "Avg.N_minus_bestfit.AICw","Avg.N_minus_bestfit.AIC","num.good.fit",
                 "0.025-CI.LT","0.975-CI.LT","0.025-CI.UT","0.975-CI.UT",
                 "percent.LT-UT>0",
                 "0.025-CI.Pl","0.975-CI-Pl","0.025-CI.Pu","0.975-CI.Pu","0.025-CI.Pl-Pu","0.975-CI.Pl-Pu")
  m<-as.data.frame(m)

  call_AICs_N<-normal_cop_AICw_matrix(r=r)
  normal_AICw<-call_AICs_N$normal_cop_AICw
  normal_AICw[normal_AICw==Inf]<-NaN
  normal_AIC<-call_AICs_N$normal_cop_AIC
  normal_AIC[normal_AIC==Inf]<-NaN

  d<-dim(r$gfc_numBS)[1]
  m$num.pairs<-(d^2)-d 
  m$`num.non-indep`<- (d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)

  bcop<-r$info_ord_copcode[,,1] # matrix with best fit copula name
  bcop[bcop==Inf]<-NaN # replacing Independent pairs with NaN
  m$`num.non-N`<-sum(bcop!=1,na.rm = T)

  r$info_ord_AIC[r$info_ord_AIC==Inf] <- NaN # replacing non-ind loc pair value by NaN
  r$info_ord_AICw[r$info_ord_AICw==Inf] <- NaN # replacing non-ind loc pair value by NaN

  bestfit.AIC<-r$info_ord_AIC[,,1]
  bestfit.AICw<-r$info_ord_AICw[,,1]

  m$Avg.N_minus_bestfit.AICw<-round(mean(normal_AICw - bestfit.AICw, na.rm = T),3)
  m$Avg.N_minus_bestfit.AIC<-round(mean(normal_AIC - bestfit.AIC, na.rm = T),3)

  pcvm<-r$gfc_p_CvM
  pcvm[pcvm==Inf]<-NaN
  pks<-r$gfc_p_KS
  pks[pks==Inf]<-NaN
  temp<-which((pcvm>0.01) & (pks>0.01),arr.ind=T)
  m$num.good.fit<-dim(temp)[1]
  t<-r$LTdep_AICw[is.finite(r$LTdep_AICw)]
  CI_LT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$`0.025-CI.LT`<-round(CI_LT[1],3)
  m$`0.975-CI.LT`<-round(CI_LT[2],3)
  t<-r$UTdep_AICw[is.finite(r$UTdep_AICw)]
  CI_UT<-unname(quantile(t,probs=c(0.025,0.975),na.rm = T))
  m$`0.025-CI.UT`<-round(CI_UT[1],3)
  m$`0.975-CI.UT`<-round(CI_UT[2],3)
  m$`percent.LT-UT>0`<-round((sum(r$LTdep_AICw-r$UTdep_AICw>0,na.rm=T)/((d^2)-d-2*(r$num_indep_loc_pair+r$num_neg_cor_loc_pair)))*100,2)
  m$`0.025-CI.Pl`<-round(rs$numericdf[5,3],3)
  m$`0.975-CI-Pl`<-round(rs$numericdf[5,4],3)
  m$`0.025-CI.Pu`<-round(rs$numericdf[6,3],3)
  m$`0.975-CI.Pu`<-round(rs$numericdf[6,4],3)
  m$`0.025-CI.Pl-Pu`<-round(rs$numericdf[10,3],3)
  m$`0.975-CI.Pl-Pu`<-list(round(rs$numericdf[10,4],3))
  #m$CI_Pl<-list(c(round(rs$numericdf[5,3],3),round(rs$numericdf[5,4],3)))
  #m$CI_Pu<-list(c(round(rs$numericdf[6,3],3),round(rs$numericdf[6,4],3)))
  #m$`CI_Pl-Pu`<-list(c(round(rs$numericdf[10,3],3),round(rs$numericdf[10,4],3)))
